Measurement of the \ce{^{64}Zn},\ce{^{47}Ti}(n,p) Cross Sections using a DD Neutron Generator for Medical Isotope Studies

### 2017-04-21 First_np_Paper/first_np_paper.tex
Merged in Shamsu's 2nd round edits

Cross sections for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} reactions have been measured for quasi-monoenergetic DD neutrons produced by the UC Berkeley High Flux Neutron Generator (HFNG).
The HFNG is a compact neutron generator designed as a \enquote{flux-trap} that maximizes the probability that a neutron will interact with a sample loaded into a specific, central location.  
The study was motivated by interest in the production of \ce{^{47}Sc} and \ce{^{64}Cu} as emerging medical isotopes.
The cross sections were measured in ratio to the \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} inelastic scattering reactions on co-irradiated indium samples.
Post-irradiation counting using an HPGe and LEPS detectors allowed for cross section determination to within 5\% uncertainty.
The \ce{^{64}Zn}(n,p)\ce{^{64}Cu} cross section for 2.765$^{+0.015}_{-0.022}$ MeV neutrons is reported as    49.315 $\pm$ 2.567 mb (relative to \ce{^{113}In}) or 46.415 $\pm$ 1.720 mb (relative to \ce{^{115}In}), and the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} cross section is reported as 26.262 $\pm$  0.822 mb.
The cross sections were determined with lower uncertainty than existing measurements and are found to be  in good agreement with both empirical and theoretical values.
This work highlights the utility of compact, flux-trap DD-based neutron sources for nuclear data measurements and potentially the production of radionuclides for medical applications.

@@

Two other potential neutron sources for (n,x) reactions exist in addition to the secondary neutron fields generated at existing isotope production facilities: reactors and neutron generators that utilize the D(T,n)$\alpha$ (\enquote{DT}) and D(D,n)\ce{^3He} (\enquote{DD}) reactions.
 While reactors produce copious quantities of neutrons, their energy spectra is often not well-suited to the preparation of high-purity samples due to the co-production of unwanted activities via neutron capture, in addition to the significant start-up costs and proliferation concerns involved in their commissioning \cite{Updegraff2013}.
 Similarly, while the higher energy 14-15 MeV neutrons produced at DT generators are capable of initiating (n,p) and (n,$\alpha$) reactions, their higher energy opens the possibility of creating unwanted activities via (n,pxn) and (n,$\alpha$xn) reactions that cannot easily be separated from the desired radionuclides.

@@

Current methodology in radiochemistry has shown recovery of upwards of 95\% of produced \ce{^{64}Cu} \cite{bhatki1969preparation,mirzadeh1992spontaneous} and \ce{^{47}Sc} \cite{Aly1971,Bokhari2010,BF02047448} from solid target designs, without the need for additional carrier.
By expanding the base of efficient reaction pathways, great advances can be made in making production of medical radionuclides more efficient and affordable for those in need.

@@

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards \cite{Capote2012,zsolnay2012technical}.
  \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations.
In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.
 After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium (HPGe) gamma-ray detector and time-dependent decay gamma-ray spectra were collected.

@@

@@

    &  & & & 0.50 \pm 0.02$^\dagger$ & 10.00 \pm 0.09$^\dagger$  & 0.248 \pm 0.005$^\dagger$ \\ \hline

  &  & & &  1.15 \pm 0.02 & 9.94 \pm 0.03 & 0.337 \pm 0.005  \\   \bottomrule
\end{tabular}%
}
\end{table*}

@@

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/scatt_angle.pdf}
 % scatt_angle.pdf: 369x295 pixel, 72dpi, 13.02x10.41 cm, bb=0 0 369 295
 \caption{Energy-angle distribution for neutrons emitted following DD fusion, for 100 keV incident deuterons.}
 \label{fig:scatt_angle}
\end{figure}

Since the samples are separated by only  8 mm from the DD reaction surface they  subtend a fairly significant  (\mytilde 17\degree)
% the above figure assumes a 5mm diameter beam seperated by 8 mm from a 10 cm diameter target, looking at the angle between the two edges
angular range in a region of impressive  (approximately $\sci{1.3}{7}$neutrons/cm$^2$s) neutron flux.
This stands in contrast to other measurements which feature collimated beams and significantly lower total neutron flux.

@@

While this shows that the sample foils experience a very narrow energy distribution of incident neutrons, an effective neutron energy window must be determined.
The MCNP6 simulation shows an identical flux-weighted average neutron energy of 2.765 MeV for both the Indium and target foils to the 1 keV level.
Due to the kinematics of DD neutron emission, $E_{max}$,  the maximum energy of a neutron subtending the target foils in this geometry is 2.783 MeV \cite{Liskien_Paulsen_1973}.
For this maximum energy, the number of reactions induced in a foil (containing $N_i$ target nuclei) is given by:

\begin{equation}
R = N_i \int_0^{E_{max}} \sigma(E) \dfrac{d\phi}{dE} dE
\end{equation}

@@

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/fracplot_new.pdf}
 % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
 \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]. The red arrows indicate the energy region that corresponds to 67\% of the total activation.}
 \label{fig:frac_plot}
\end{figure}

@@

The spectra for the different samples were summed and the net peak areas were fitted using gf3, part of the RadWare analysis package from Oak Ridge National Laboratory   \cite{radford2000radware, Radford1995}.
The background-subtracted integrated counts in each photopeak, as well as the counting duration for each experiment, are tabulated in \autoref{tab:peak_counts}.

@@

Target Foil Mass (g)               & 0.538         & 0.451       & 0.452 & 0.337     & 0.337   \\ 
Irradiation Time, $t_1$ (s)           & 10800         & 10800       & 12629 & 11837     & 14254   \\ 
Start of Count, $t_2$ (s)             & 12585         & 26985       & 14919 & 101245    & 16644   \\ 
End of Count, $t_3$ (s)               & 103773        & 80993       & 68919 & 187669    & 110275   \\ 
Photopeak Counts, 336 keV (\ce{^{115m}In}) & 113665 $\pm$ 1490 & 74321 $\pm$ 275 & 39895 $\pm$ 201 & 2122 $\pm$ 55 & 55102 $\pm$ 268   \\ 
Photopeak Counts, 391 keV (\ce{^{113m}In}) & 3382 $\pm$ 171    & 860 $\pm$ 40    & 2545 $\pm$ 54 &    \hrulefill       &  \hrulefill      \\ 
Photopeak Counts, 511 keV (\ce{^{64}Cu})   & 16055 $\pm$ 643   & 12852 $\pm$ 118 & 24464 $\pm$ 159 &    \hrulefill       &   \hrulefill     \\ 

@@

If this irradiation lasts for  some time $t_1$, and gamma ray spectrum acquisition begins at a later time $t_2$ and ends at  $t_3$, then the number of product decays ($D$; with decay constant $\lambda$) during the acquisition will be:

\begin{equation}
\begin{aligned}
D &= \dfrac{R}{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}}
\\
&= \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}} 
\end{aligned}
\end{equation}

If this decay emits a gamma ray with absolute intensity $I_\gamma$ (photons emitted per decay), and is detected with an absolute efficiency  of $\epsilon_\gamma$ (photons detected / photons emitted), then the number of observed decays between $t_2$ and $t_3$ will be:

@@

\begin{align}
N_{obs} &= D \epsilon_\gamma I_\gamma \\
&=  \epsilon_\gamma I_\gamma  \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}} \nonumber
\end{align}

@@

\begin{equation}\label{eqn:single_xs_eqn}
\sigma\pp{\bar{E}} = \dfrac{N_{obs}\lambda}{N_T \epsilon_\gamma I_\gamma  \phi\pp{\bar{E}}  \pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}}}
\end{equation}

@@

\begin{align}\label{eqn:calc_eqn}
\dfrac{\sigma_P}{\sigma_{In}} &=  \dfrac{A_P}{A_{In}} \dfrac{N_{T,In}}{N_{T,P}} \dfrac{\lambda_P}{\lambda_{In}} \dfrac{e^{\lambda_{In}t_1}-1}{e^{\lambda_{P}t_1}-1} \times \\
&\times \dfrac{e^{-\lambda_{In}t_3}-e^{-\lambda_{In}t_2}}{e^{-\lambda_{P}t_3} - e^{-\lambda_{P}t_2}} \dfrac{\epsilon_{In}}{\epsilon_P}  \dfrac{I_{\gamma,In}}{I_{\gamma,P}} \dfrac{e^{-\mu_{In}x_{In}/2}\times e^{-\mu_{In}x_{P}}}{e^{-\mu_{P}x_{P}/2}} \nonumber
\end{align}

@@

\item $A$ is the integrated counts under a photopeak [counts],

@@

\item $t_1$ is the irradiation time [s],

\item $t_2$ is the time between the start-of-beam and the start of counting [s],

\item $t_3$ is the time  between the start-of-beam and the end of counting [s],

@@

This total  uncertainty will be plotted as the cross section error bar in the excitation function for these production reactions.

@@

\comment{Karl:  \enquote{For the figures, let me suggest you make a red solid rectangle whose dimensions represent the error bars in both energy and cross section.  Then put a large red circle around it, whose diameter is of order 10\% of the size of the whole graph, to focus the eyes.  In the caption, mention that the size of the rectangle represents the error limits.}  Lee: \enquote{I would suggest that you zoom in on the energy and cross section value we measured (say 2.5-3.0 MeV and 10-70 mb).  Use a thicker line for our data point and/or make it larger. - Lee}}

@@

\begin{equation}
A = \lambda N_{product} = \eta \phi \sigma N_{target}
\end{equation}

>>>>

Cross sections for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} reactions have been measured for quasi-monoenergetic DD neutrons produced by the UC Berkeley High Flux Neutron Generator (HFNG).
The HFNG is a compact neutron generator designed as a \enquote{flux-trap} that maximizes the probability that a neutron will interact with a sample loaded into a specific, central location.  
The study was motivated by interest in the production of \ce{^{47}Sc} and \ce{^{64}Cu} as emerging medical isotopes.
The cross sections were measured in ratio to the \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} inelastic scattering reactions on co-irradiated indium samples.
Post-irradiation counting using an HPGe and LEPS detectors allowed for cross section determination to within 5\% uncertainty.
The \ce{^{64}Zn}(n,p)\ce{^{64}Cu} cross section for 2.765$^{+0.015}_{-0.022}$ MeV neutrons is reported as    49.315 $\pm$ 2.567 mb (relative to \ce{^{113}In}) or 46.415 $\pm$ 1.720 mb (relative to \ce{^{115}In}), and the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} cross section is reported as 26.262 $\pm$  0.822 mb.
The measured cross sections  are found to be  in good agreement with existing measured values but with lower uncertainty (\textless 5\%), and also in agreement with  theoretical values.
This work highlights the utility of compact, flux-trap DD-based neutron sources for nuclear data measurements and potentially the production of radionuclides for medical applications.

@@

Two other potential neutron sources for (n,x) reactions exist in addition to the secondary neutron fields generated at existing isotope production facilities: reactors and neutron generators that utilize the D(T,n)$\alpha$ (\enquote{DT}) and D(D,n)\ce{^3He} (\enquote{DD}) reactions.
 While reactors produce copious quantities of neutrons, their energy spectra is often not well-suited to the preparation of high-purity samples due to the co-production of unwanted activities via neutron capture, in addition to the significant start-up costs and proliferation concerns involved in their commissioning \cite{Updegraff2013}.
 Similarly, while the higher energy 14-15 MeV neutrons produced at DT generators are capable of initiating (n,p) and (n,$\alpha$) reactions, their higher energy opens the possibility of creating unwanted activities via (n,pxn) and (n,$\alpha$xn) reactions that cannot easily be separated from the desired radionuclides. DT generators may also often be limited by the restricted use of tritium at many institutions.

@@

Current methodology in radiochemistry has shown recovery of upwards of 95\% of produced \ce{^{64}Cu} \cite{bhatki1969preparation,mirzadeh1992spontaneous} and \ce{^{47}Sc} \cite{Aly1971,Bokhari2010,BF02047448} from solid target designs, without the need for additional carrier.
By expanding the base of efficient reaction pathways, great advances are possible in making production of medical radionuclides more efficient and affordable for those in need.  It is this desire to improve the options available for modern medical imaging and cancer therapy which has motivated the campaign of nuclear data measurements for isotope production at the UC Berkeley HFNG.

@@

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards \cite{Capote2012,zsolnay2012technical}.
  \autoref{tab:foil_specs} lists physical characteristics of each foil for the various irradiations.
In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.
 After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium (HPGe) gamma-ray detector and time-dependent decay gamma-ray spectra were collected.

@@

  &  & & &  1.15 \pm 0.02 & 9.94 \pm 0.03 & 0.337 \pm 0.005  \\  \hline

@@

    &  & & & 0.50 \pm 0.02$^\dagger$ & 10.00 \pm 0.09$^\dagger$  & 0.248 \pm 0.005$^\dagger$ \\ \bottomrule

\end{tabular}%
}
\end{table*}

@@

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/scatt_angle.pdf}
 % scatt_angle.pdf: 369x295 pixel, 72dpi, 13.02x10.41 cm, bb=0 0 369 295
 \caption{Energy-angle distribution for neutrons emitted following DD fusion, for 100 keV incident deuterons \cite{Liskien_Paulsen_1973}.}
 \label{fig:scatt_angle}
\end{figure}

Since the samples are separated by only  8 mm from the DD reaction surface they  subtend a fairly significant  (\mytilde 17\degree)
% the above figure assumes a 5mm diameter beam seperated by 8 mm from a 10 cm diameter target, looking at the angle between the two edges
angular range in a region of  high  (approximately $\sci{1.3}{7}$neutrons/cm$^2$s) neutron flux.
This stands in contrast to other measurements which feature collimated beams and significantly lower total neutron flux.

@@

While this shows that the sample foils experience a very narrow energy distribution of incident neutrons, an effective neutron energy window must be determined.
The MCNP6 simulation shows an identical flux-weighted average neutron energy of 2.765 MeV for both the Indium and target foils to the 1 keV level.
Due to the kinematics of DD neutron emission, $E_{max}$,  the maximum energy of a neutron subtending the target foils in this geometry is 2.783 MeV \cite{Liskien_Paulsen_1973}.
For this maximum energy, the number of reactions induced in a foil (containing $N_T$ target nuclei) is given by:

\begin{equation}
R = N_T \int_0^{E_{max}} \sigma(E) \dfrac{d\phi}{dE} dE
\end{equation}

@@

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/fracplot_new.pdf}
 % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
 \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]. The solid red boundaries indicate the energy region that corresponds to 68.2\% of the total activation.}
 \label{fig:frac_plot}
\end{figure}

@@

The spectra for each sample were summed and the net peak areas were fitted using gf3, part of the RadWare analysis package from Oak Ridge National Laboratory   \cite{radford2000radware, Radford1995}.
The background-subtracted integrated counts in each photopeak, as well as the counting duration for each experiment, are tabulated in \autoref{tab:peak_counts}.

@@

Target Foil Mass (g)               & 0.538         & 0.521       & 0.542 & 0.337     & 0.337   \\ 
Irradiation Time, $t_i$ (s)           & 10800         & 10800       & 12629 & 11837     & 14254   \\ 
Delay Time, $t_d$ (s)             & 1785         & 16185       & 2290 & 89408    & 2390   \\ 
Counting Time, $t_c$ (s)               & 91188        & 54008       & 54002 & 86424    & 93631   \\ 
Photopeak Counts, 336 keV (\ce{^{115m}In}) & 113665 $\pm$ 1490 & 76321 $\pm$ 275 & 39895 $\pm$ 201 & 2122 $\pm$ 55 & 55102 $\pm$ 268   \\ 
Photopeak Counts, 391 keV (\ce{^{113m}In}) & 3382 $\pm$ 171    & 890 $\pm$ 40    & 3505 $\pm$ 54 &    \hrulefill       &  \hrulefill      \\ 
Photopeak Counts, 511 keV (\ce{^{64}Cu})   & 16055 $\pm$ 643   & 12852 $\pm$ 118 & 27164 $\pm$ 159 &    \hrulefill       &   \hrulefill     \\ 

@@

If the target is subjected this flux  for an irradiation  time $t_i$ and decays for a  delay time $t_d$ (after end-of-beam) before gamma ray spectrum acquisition occurs for a counting time  $t_c$, then the number of product decays ($N_D$; with decay constant $\lambda$) during the acquisition will be:

\begin{align}
N_D &= \dfrac{R}{\lambda}\pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}}
\\
&= \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}} \nonumber
\end{align}

% \begin{equation}
% \begin{aligned}
% N_D &= \dfrac{R}{\lambda}\pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}}
% \\
% &= \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}} 
% \end{aligned}
% \end{equation}

If this decay emits a gamma ray with absolute intensity $I_\gamma$ (photons emitted per decay), and is detected with an absolute efficiency  of $\epsilon_\gamma$ (photons detected / photons emitted), then the number of observed gamma rays during the acquisition will be:

@@

\begin{align}
N_{\gamma} &= N_D \epsilon_\gamma I_\gamma \\
&=  \epsilon_\gamma I_\gamma  \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{1 - e^{-\lambda t_i}}  e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}} \nonumber
\end{align}

@@

\begin{align}\label{eqn:single_xs_eqn}
\sigma\pp{\bar{E}} = \dfrac{N_{\gamma}\lambda}{N_T \epsilon_\gamma I_\gamma  \phi\pp{\bar{E}}  \pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}}}
\end{align}

% \begin{equation}\label{eqn:single_xs_eqn}
% \sigma\pp{\bar{E}} = \dfrac{N_{obs}\lambda}{N_T \epsilon_\gamma I_\gamma  \phi\pp{\bar{E}}  \pp{1 - e^{-\lambda t_i}} e^{-\lambda t_d} \pp{1 - e^{-\lambda t_c}}}
% \end{equation}

@@

\begin{align}\label{eqn:calc_eqn}
\dfrac{\sigma_P}{\sigma_{In}} &=  \dfrac{N_{\gamma,P}}{N_{\gamma,In}}  \dfrac{N_{T,In}}{N_{T,P}} \dfrac{\lambda_P}{\lambda_{In}} \pp{\dfrac{1 - e^{-\lambda_{In}t_i}}{1 - e^{-\lambda_{P}t_i}}} \dfrac{e^{-\lambda_{In}t_d}}{ e^{-\lambda_{P}t_d}} \times \\
&\times \pp{\dfrac{1 - e^{-\lambda_{In}t_c}}{1 - e^{-\lambda_{P}t_c}}} \dfrac{\epsilon_{In}}{\epsilon_P}  \dfrac{I_{\gamma,In}}{I_{\gamma,P}} \dfrac{e^{-\mu_{In}x_{In}/2}\times e^{-\mu_{In}x_{P}}}{e^{-\mu_{P}x_{P}/2}} \nonumber
\end{align}

@@

% \item $N_{\gamma}$ is the integrated counts under a photopeak [counts],
\item $N_{\gamma}$ is the integrated counts under a photopeak,

@@

\item $t_i$ is the irradiation time [s],

\item $t_d$ is the delay time (between the end-of-beam and the start of counting) [s],

\item $t_c$ is the counting time   [s],

@@

This total  uncertainty is plotted as the cross section uncertainty in \autoref{fig:sc_xs} and \autoref{fig:cu_xs_zoom}.

@@

% \comment{Karl:  \enquote{For the figures, let me suggest you make a red solid rectangle whose dimensions represent the error bars in both energy and cross section.  Then put a large red circle around it, whose diameter is of order 10\% of the size of the whole graph, to focus the eyes.  In the caption, mention that the size of the rectangle represents the error limits.}  Lee: \enquote{I would suggest that you zoom in on the energy and cross section value we measured (say 2.5-3.0 MeV and 10-70 mb).  Use a thicker line for our data point and/or make it larger. - Lee}}

@@

\begin{equation}
A = \lambda N_{product} = \eta \phi \sigma N_{T}
\end{equation}
